---
title: "Cleaning data"
author: "Dien Giau (Richard) Bui"
date: "2/7/2022"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library
```{r warning=FALSE, message=FALSE, include=TRUE}
library(tidyverse) # for MacOS: library(dplyr)
library(stringr)
library(janitor)
library(tidyr)
library(skimr)
library(readxl)
```


## What is cleaning data?

- Data is generally "dirty": it has a lot of errors/problems
    - Weird column names
    - Not in tidy format
    - Wrong data type: a numeric variable but is stored as characters
    - Some typos in data: "femela" (should be "female")
    - Duplicates: some rows are repeated
    - Missing data: some rows/columns have no data

- "Garbage in, garbage out"
- So, cleaning step makes sure we have a clean data before doing any statistical analysis
- That's why we start this course by a lecture on cleaning data

## Our data in today lecture

- The US tuition fee in different states and territories
- A picture on the data: [click here](https://onlinembapage.com/wp-content/uploads/2016/03/AverageTuition_Part1b.jpg)
- I borrowed the data from: [Tidy Tuesday project](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018)
- I downloaded data and made it messy/dirty and store it in our "data/raw/us_avg_tuition.xlsx"
- Let's clean this data
    - I noted step by step in this slide

## 1. Load/import data to R

- Before load/import data, let's open it by MS Excel
- Load data to R by two ways:
    - By RStudio using mouse click
    - By code
    
## Import data

```{r}
Tuition = read_excel("data/raw/us_avg_tuition.xlsx")
# View(Tuition)
```


## Discussion [1]

- By hover the mouse over columns, we find that the data type is imported incorrectly
    - It is numeric data, but now is imported as character
    
- Columns names: quite non-standard
```{r}
names(Tuition)
```

- For example, can you try to take one column for the year "2004-2005":
```{r eval=FALSE}
Tuition$2004-05
```

## Discussion [2]

- One row is duplicate/repeated: can you help me to find which row?
- One row has no data (missing data)
- Some state names are in abbreviation, not with full name: can you find them?

- **Conclusion:** there are a lot of problems with data, let's clean it in next slides

## 2. Clean column names

```{r}
Tuition = clean_names(Tuition)
head(Tuition[,1:3])
```

## 3. Data type casting

- Character is not the same as numeric, for example
```{r eval=FALSE}
sqrt("16") # we can't apply numeric method to a character
sqrt(16)
```


- We want to convert tuition from character type to numeric
- In R, we will use `as.numeric` function:

```{r}
as.numeric("16")
```

## Convert character to numeric columns

It is a bit long to type all of the below code, but we will return on how to make it shorter in a later lecture.

```{r warning=FALSE}
Tuition = Tuition %>% 
  mutate(
    x2004_05 = as.numeric(x2004_05),
    x2005_06 = as.numeric(x2005_06),
    x2006_07 = as.numeric(x2006_07),
    x2007_08 = as.numeric(x2007_08),
    x2008_09 = as.numeric(x2008_09),
    x2009_10 = as.numeric(x2009_10),
    x2010_11 = as.numeric(x2010_11),
    x2011_12 = as.numeric(x2011_12),
    x2012_13 = as.numeric(x2012_13),
    x2013_14 = as.numeric(x2013_14),
    x2014_15 = as.numeric(x2014_15),
    x2015_16 = as.numeric(x2015_16)
  )
```


## Note on `pipe` or `%>%`

- Traditionally, we will write function like:
```{r eval=FALSE}
mutate(Tuition, ...)
```


- But in the previous example, I wrote:

```{r eval=FALSE}
Tuition %>% mutate(...)
```

- In plain English:
    1. We take `Tuition` data
    2. Then `mutate` to add/update new columns to the data
    3. and so on, we can continue forever with pipe

- It makes our code easier to read and follow, so I will use it a lot in this class

## 4. Remove duplicates

- Check duplicate by `duplicated`

```{r}
duplicated(Tuition)
```

- Remove duplicates by `distinct`
```{r}
Tuition = Tuition %>% 
  distinct()
```


## 5. Remove missing data

- Missing data is everywhere
- We have several ways to deal with missing:
    - Find out data
    - Fill by a make-sense number: fill by zero, by industry average, ...
    - Or remove missing rows
- We will remove missing for convenience in our data
    - Can you guess how we can remove the missing row?

## Remove missing by filter out
```{r}
Tuition = Tuition %>% 
  filter(state != "Puerto Rico")
```


## 6. String manipulation in `state` columns

- Recall that:
    - Some state names are in abbreviation, not with full name: "CA", "NY"
- We use package "`stringr`" and its function "`str_replace`"
```{r}
Tuition = Tuition %>% 
  mutate(
    state = str_replace(state, "CA", "California")
  ) %>% 
  mutate(
    state = str_replace(state, "NY", "New York")
  )
```

## 7. Wide to long data

- We can transform data from wide to long as following:
```{r}
Tuition = Tuition %>% 
  pivot_longer(-state, 
               names_to = "year", 
               values_to = "tuition")
head(Tuition)
```

## Quiz

- Can you transform the `year` column to be numeric from 2004 to 2015?
- For example: `x2004_05` will become `2004`, `x2005_06` will become `2005`

## Transform year

- Will solve in class
```{r}
# add code here
```

## Look how clean of the final data
```{r}
summary(Tuition)
```

## Conclusion

- So many steps to clean a data
- Need to document all process in an Rmd file like this
    - or you can copy each step's code + results to MS `docx` file, but will need more efforts
- By the end, you can run everything together to get a report: pdf, word, html format
- Final tip: save this cleaned data to a `data/process` folder

## Save data
```{r}
saveRDS(Tuition, "data/process/Tuition_clean.rds")
```

