---
title: "Test 01: One variable of interest Y"
author: "Richard"
date: "1/29/2022"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r warning=FALSE, message=FALSE}
library(tidyverse) 
# if you're using macOS: library(dplyr)
library(skimr)
```

## One variable of interest

- Sometimes, we often analyze/care about only one variable: salary, gender, interest, returns, score, pass/fail an exam, ...
- We can classify these measures into two groups:
    - Numeric variables: which ones above are numeric?
    - Categorical variables: which ones?
- Thanks to their data structure, we need different statistical tests applied to them when we ask questions about them
- Let check one example with a categorical variable: gender


## Prepare Data

Please read the intro about data at [here](https://www.openintro.org/data/index.php?data=hsb2)

```{r}
Hsb <- within(
  read.csv("https://stats.idre.ucla.edu/stat/data/hsb2.csv"), {
    race <- as.factor(race)
    schtyp <- as.factor(schtyp)
    prog <- as.factor(prog)
})
```

## Check data

```{r}
head(Hsb[,1:8])
```

## Summary data

```{r}
summary(Hsb)
# skim_without_charts(Hsb)
```

## Questions

- How the authors construct `female` variable in the `Hsb` dataset.
- How much is the female ratio (female/total students)?
    - We often compare this ratio to which number/ratio?

## Binomial test

- Hypothesis: Does the female ratio is equal to 0.5 or 50%?
- Null hypothesis H0: The female ratio is 50%.
- Alternative hypothesis H1: The female ratio is different from 50%.

## R function

- Function: `prop.test`
- Usage: `prop.test(x, n, p)`
    - Recall: how to read help documentation in R?
    - `?prop.test`

## R code example

```{r}
prop.test(sum(Hsb$female), length(Hsb$female), p = 0.5)
```

## Discussion

- In the sample, the female ratio is 54.5%
- The binomial test shows that the p-value is 0.2293, which is larger than the significance level (e.g., 10%)
- So we can't reject the null hypothesis
    - we don't have enough evidence to conclude that the female ratio is different than 50%

## Similar questions in our life

- We often have same questions in our daily life
    - Does wearing mask prevent us from covid virus?
    - Does wearing helmet will help motorcylists have less serious traffic accidents?

## Short quizes in classes

I will give you a sample data, let apply the `prop.test` with that data in class.

## Numeric variable

- Next, switching to a numeric variable, which can receive any continuous value:
    - e.g., salary, returns, interest rate, ...
    
- We often want to know the mean (centralized tendency) and the variance/standard deviation of this variable:
    - e.g.1., what is the average salary after we graduated and got the first job
    - e.g.2., what is the average `write` score of all students in the class

## Questions

- Check again, what is average `write` score in our `Hbs` data
- Is it equal to 50, or different

## t test

- Hypothesis: Does the write score is equal to 50?
- Null hypothesis H0: The write score is 50.
- Alternative hypothesis H1: The write score is different from 50.

## R function

- Function: `t.test`
- Usage:

## R code example

```{r}
t.test(Hsb$write, mu = 50)
```

## Discussion

- How to read the results?

## Short quizes again

- Let apply `t.test` more

## Another important stat to measure central tendency: median

- Let do the same test, but ask if the median is equal to 50 or not

```{r}
wilcox.test(Hsb$write, mu = 50)
```



## Next lecture

- This lecture introduces to you three important tests:
    - `prop.test`
    - one-sample `t.test`
    - one-sample median test `wilcox.test`
- We will consider two variables at the same time:
    - One is our variable of interest such as `write` score
    - Another one is another factor: such as `female` and `male` students
    - So the question is more like: are the write scores the same between female and male students? Or female students write better (thanks to their gifted writing skills) so they score higher?
    
- Let's see in the next lecture

