---
title: 'Test 02: Numeric Y vs categorical X'
author: "Richard"
date: "1/30/2022"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r warning=FALSE, message=FALSE}
library(tidyverse) # if you're using macOS, you can run: library(dplyr)
library(skimr)
```

## Prepare Data

Please read the intro about data at [here](https://www.openintro.org/data/index.php?data=hsb2)

```{r}
Hsb <- within(
  read.csv("https://stats.idre.ucla.edu/stat/data/hsb2.csv"), {
    race <- as.factor(race)
    schtyp <- as.factor(schtyp)
    prog <- as.factor(prog)
})
```

## Recall

- Numeric variable
- Categorical variable
- What is the key difference between them?

## Some questions between numeric Y and categorical X

- We care about numeric Y for different groups in categorical X
- For example: Y is salary/score, X is gender
    - Do male employees earn more than the female co-workers
    - Do female students have higher write score than the male friends

## R function

- Function: `t.test`
- Usage: `t.test(y ~ x)
- `y` is a numeric variable
- `x` is a categorical variable with two groups
    - e.g., `female` includes only two values 0 and 1

## R code example

```{r}
t.test(Hsb$write ~ Hsb$female)
```

## Assumption of `t.test`

- It requires `y` to be normal distributed!
    - it is a very strong assumption
- If `y` is not normal distributed, we aren't confident to use t-test to answer the above research question
- So, let's check the normality assumption of `y`
    - or, we ask if `write` is normal distributed first, before we use the t-test
    
## Check normality of `write` by histogram

```{r}
hist(Hsb$write)
```

## A normality test

```{r}
shapiro.test(Hsb$write)
```

## Discuss

- It seems that `write` does not follow normal distribution
- So we can't use t-test in this case
- Do we have an alternative test, when we don't have the normality assumption
    - Yeah! We can use The Wilcoxon-Mann-Whitney test

## R function

- Function: `wilcox.test`
- Usage: `wilcox.test(y ~ x)`
- `y` is a numeric variable
- `x` is a categorical variable with two groups
    - e.g., `female` includes only two values 0 and 1

## R code example

```{r}
wilcox.test(Hsb$write ~ Hsb$female)
```

## Extention: more than 2 groups

- `t.test` works for 2-group X only
- If we want to check mean difference for more than 2 groups, we need to use one-way ANOVA 
- For example: if the `write` score is the same for every program

```{r}
Hsb %>% count(prog)
```

## R code example
```{r}
summary(aov(Hsb$write ~ Hsb$prog))
```

## Discussion

- A very small p-value indicates that the write score is not the same across programs

## Short quiz

- Rewrite ANOVA code above using `%>%` instead of `summary(aov(...))`

## In sum

- In this lecture, we learn:
    - Two-sample `t.test`
    - Wilcoxon-Mann-Whitney test `wilcox.test`
    - ANOVA `aov`

- Next lecture goes to a more common case when both Y and X are numeric variables






