---
title: 'Test 03: Categorical Y vs categorical X'
author: "Richard"
date: "1/30/2022"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r warning=FALSE}
library(tidyverse) 
# if you're using macOS, you can run: library(dplyr)
library(skimr)
```



## Prepare Data

```{r}
Hsb <- within(
  read.csv("https://stats.idre.ucla.edu/stat/data/hsb2.csv"), {
    race <- as.factor(race)
    schtyp <- as.factor(schtyp)
    prog <- as.factor(prog)
})
```

## Categorical Y and categorical X

 For example, we use the hsb2 data file, let’s see if there is a relationship between the type of school attended (schtyp) and students’ gender (female).

## Frequency table

- The below code show the frequency table between gender and school type:

```{r}
Hsb %>% count(female, schtyp)
```

- Can you explain this data?

## A matrix format

- Each row is gender
- Each column is each school type

```{r}
table(Hsb$female, Hsb$schtyp)
```


## Chi-square test

```{r}
chisq.test(table(Hsb$female, Hsb$schtyp))
```

## Discussion

- These results indicate that there is no statistically significant relationship between the type of school attended and gender (chi-square with one degree of freedom = 0.0005, p = 0.98).


## Another example

Please check the relationship between gender (`female`) and socio-economic status (`ses`) using chi-square test.

## Assumption of chi-square test

- Remember that the chi-square test assumes the expected value of each cell is five or higher.
- However, if this assumption is not met in your data, please see the section on Fisher’s exact test below.

## Fisher's exact test

```{r}
fisher.test(table(Hsb$female, Hsb$schtyp))
```

## Practice

Please check the relationship between race (`race`) and school type (`schtyp`)

## Quiz time

Hmm...