---
title: "07 Don't repeat yourself"
author: "Dien Giau (Richard) Bui"
date: "4/20/2022"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 1, digits = 2) #set to two decimal 
```

## Load library

```{r warning=FALSE}
library(dplyr)
library(skimr)
library(stringr)
library(broom)
```

```{r include=FALSE}
analyze_subsize = function(size_type){
  ## YOU CONTINUE HERE
  Qs_subset = Qs %>% filter(size==size_type)
  print(mean(Qs_subset$size))
  m2 = lm(score ~ international_students + faculty_count + student_faculty_ratio, data=Qs_subset)
  return(tidy(m2))
}
```

## Introduction

- Life is full of repeated routine
    - Repeat calculation or analysis
- This note provides some techniques to iterate/repeat things in a programming way such as:
    - Define your own function to repeat things
    - Loop or mapping
    
## Example

- We may ask if how much is the ranking score of universities across different size
- and for each group size, if the number of international students, faculties, and student-faculty-ratio can affect the `score`
- Load data:

```{r message=FALSE, warning=FALSE}
data_path = "data/raw/qs-world-university-rankings-2017-to-2022-V2.csv"
Qs = readr::read_csv(data_path)

Qs = Qs %>% 
  mutate(research_output = str_replace(research_output, "Very high", "Very High"))
Qs = Qs %>% filter(!is.na(research_output))
Qs = Qs %>% filter(!is.na(type))
Qs = Qs %>% 
  mutate(private = ifelse(type=="Private", 1, 0))
```

## How many size groups?

```{r}
Qs %>% count(size)
```

## First, focus on L size

```{r}
Qs_L = Qs %>% filter(size == "L")
mean(Qs_L$score, na.rm = TRUE)
m1 = lm(score ~ international_students + faculty_count + 
          student_faculty_ratio, data=Qs_L)
tidy(m1)
```

## Then, we continue with M size

```{r}
Qs_M = Qs %>% filter(size == "M")
mean(Qs_M$score, na.rm = TRUE)
m2 = lm(score ~ international_students + faculty_count + 
          student_faculty_ratio, data=Qs_M)
tidy(m2)
```

## Discussion

- The process to analyze for L and M groups is very similar
    - Filter: one filter "L" and one filter "M"
    - Take mean of `score`: same
    - Run a regression: same formula
    - Clean/tidy the model coefficients: same

- So, actually, they are only different at one input in filter data in the first step

## Define own function

- So, we can think about create your own function so that we can repeat this process easily
- A function needs to have:
    - Input: input of the function, in our case, should be the value to filter size, it could receive "L", "M", "S", "XL"
    - Process with this input
    - Return output by the end
- Look like this:
```{r eval=FALSE}
function_name = function(Input){
  # PROCESS DATA HERE to get Output
  
  # RETURN OUTPUT HERE
  return(Output)
}
```


## For example

- Write a function to take square of a number

```{r}
# step 1: define your function
calculate_square = function(x){
  x_square = x*x
  return(x_square)
}
# step 2: use/apply the function
calculate_square(4)
```


## Try to write your function here

- AN EMPTY SLIDE HERE WITH INTENTION
```{r eval = FALSE}
analyze_subsize = function(size_type){
  ## YOU CONTINUE HERE
}
```


## Apply the function for each group

```{r message=FALSE, warning=FALSE}
analyze_subsize(size_type = "L")
#analyze_subsize(size_type = "M")
#analyze_subsize(size_type = "S")
#analyze_subsize(size_type = "XL")
```

## Discussion

- After having a magic function, we only need to use/apply the function 4 times
    - And if we need later, we also just simply apply more, no need to re-write the whole long process
- If we add more analysis (t-test or so), we simply fix the source code of function
    - So very easy to maintain the analysis process

## But ... 

- Do you notice that we aim to avoid repeating things in programming
- So even we now only have a magic function, we still need to repeat 4 times (and even more, e.g., 1,000 groups)

```{r}
#analyze_subsize(size_type = "L")
#analyze_subsize(size_type = "M")
#analyze_subsize(size_type = "S")
#analyze_subsize(size_type = "XL")
```

## How to avoid this repeatation?

- In R, we can use some loop/mapping to avoid this kind of repeat
- Let me show you with loop then with mapping later

## Use loop to avoid repeat a function many times

```{r message=FALSE, warning=FALSE}
all_size_type = c("M", "L", "S", "XL")
output = vector(mode = "list", length = 4)
for (i in 1:4){
  output[[i]] = analyze_subsize(size_type = all_size_type[i])
}
```

## See our output vector

```{r}
# for L size
output[[2]]
```

## Another way is mapping or apply

- `lapply` is apply and store output as a list
```{r, message=FALSE, warning=FALSE}
output = lapply(X = all_size_type, FUN = analyze_subsize)
# for L size
output[[2]]
```


## Thank you

- Thank you for today
- We will have more practices in class


















